Modelling the emergence of whisker barrels

### 2020-06-01 paper/briefpaper/brief.tex
Figure name refactoring
More text for paper

\begin{figure}
  \begin{fullwidth}
    %\includegraphics[width=\linewidth]{./MainFig.png}
    \includegraphics[width=\linewidth]{./MainFig_superlowres.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
      simulation approaching that obtained from barrels in \textbf{A} (dotted
      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta$ (the sum of differences in
      area; units \cmnt{mm$^2$});
      \cmnt{grey hexagons show the measure of localization, $\omega =
        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
      \textbf{E} \cmnt{The localization measure plotted across the domain
        shows that the structure in the connection density which has developed
        by $t=50000$ is largely absent at $t=1000$. The
        colour map
        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

A cortical boundary enclosing 41 corresponding barrels was traced from a
cytochrome oxidase stain from \cite{zheng_signal_2001}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes. From random initial conditions for
$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
within \cmnt{$\approx 10000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
in shape between real and simulated barrels revealed a strong correspondence
(see Fig.\,\ref{fig:main}D).

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./paramsweep.png}
    \caption{\cmnt{\textbf{A} Parameter sweep caption goes here}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\cmnt{Parameter sweep paragraph goes here.}

@@

To further investigate the interplay of genetic and intrinsic mechanisms we
simulated a seminal barrel duplication experiment
\citep{shimogori_fibroblast_2005}, first by solving on a cortical domain
comprising two separate, identically shaped cortical boundaries. In this case
one distinct cortical field per thalamic projection emerged, within one of the
two boundaries only, i.e., no duplication. However, merging the two fields to
create an extended mirror-symmetric boundary shape, with an anterior-posterior
guidance field reflected at the join, gave rise to two mirror-symmetrical
barrel fields comprising 2N barrels, i.e., two cortical fields for each
thalamic projection (Fig.\,\ref{fig:main}E). Together these results suggest that by the
misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
effectively created one large barrel field rather than two distinct cortical
fields.

@@

The present results suggest that the key requirements for the emergence of
realistic barrel patterning are i) at each cortical location thalamocortical
projections compete for a limited number of available synaptic connections
(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
given projection is reduced by the density of other projections
(Eq.\,\ref{eq:comp}), and iii) \cmnt{REVIEW: No longer true. Try plotting $a$
  vs $t$} the branch density of each projection is conserved over time.

>>>>

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta$ (the sum of differences in
      area; units \cmnt{mm$^2$});
      \cmnt{grey hexagons show the measure of localization, $\omega =
        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
      \textbf{E} \cmnt{The localization measure plotted across the domain
        shows that the structure in the connection density which has developed
        by $t=30000$ is largely absent at $t=1000$. The
        colour map
        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

A cortical boundary enclosing 41 corresponding barrels was traced from a
cytochrome oxidase stain from \cite{zheng_signal_2001}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes. From random initial conditions for
$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
within \cmnt{$\approx 20000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
(see Fig.\,\ref{fig:main}D). \cmnt{In a chosen region, the ratio of the connection density
  of the TC type $i$ which makes the densest connections to the sum of all
  connection densities gives a measure of the localization of the connection
density. We computed this measure across the entire barrel domain and named it $\omega$
(Fig.\,\ref{fig:main}D). $\omega$ increases during the first $20000$ simulation
steps and converges on a stable value. Fig.\,{fig:main}E shows
the localization measure plotted as a luminance map, giving a more graded
impression of the structure which develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\cmnt{\textbf{A} Parameter sweep caption goes here}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\cmnt{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
  iterations plotted as colour maps for $180$ parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows
  results for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$ varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
  are yellow. Lower values of the difference between real and simulated barrel
  area indicates good pattern matching. Higher values of $\omega$ indicate
  better localization and are coloured yellow. White squares indicate that the
  simulation was not stable for that parameter set.}

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the maximally
  localized solution, which for $D$ should be small. Thus, there is a
  trade-off between good localization and good Voroni pattern conformation. We
  chose the parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A to present in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii) and iii) indicate the effect
  of varying $D$ in either direction from its optimum. ii) is a less
  exaggerated version of the pattern in i), whereas iii) shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

To further investigate the interplay of genetic and intrinsic mechanisms we
simulated a seminal barrel duplication experiment
\citep{shimogori_fibroblast_2005}, first by solving on a cortical domain
comprising two separate, identically shaped cortical boundaries. In this case
one distinct cortical field per thalamic projection emerged, within one of the
two boundaries only, i.e., no duplication. However, merging the two fields to
create an extended mirror-symmetric boundary shape, with an anterior-posterior
guidance field reflected at the join, gave rise to two mirror-symmetrical
barrel fields comprising 2N barrels, i.e., two cortical fields for each
thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
effectively created one large barrel field rather than two distinct cortical
fields.

@@

The present results suggest that the key requirements for the emergence of
realistic barrel patterning are i) at each cortical location thalamocortical
projections compete for a limited number of available synaptic connections
(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
given projection is reduced by the density of other projections
(Eq.\,\ref{eq:comp}), and iii) the branch density of each projection is
conserved over time.
